---
title: " Asteroid Data Analysis"
author: "Issac Chan, Lance Yuan, and Eva Wang"
format:
  html:
    theme: solar
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r, message=FALSE}
# Import data
total_data <- read.csv('nasa.csv', header = T, stringsAsFactors = T)

# Library Everything
library(ggplot2)
library(reshape) 
library(ggcorrplot)
library(splitstackshape) 
library(randomForest)
library(xgboost)
library(caret)
library(OptimalCutpoints) 
library(ggforce)
library(glmnet)
library(neuralnet)
library(pROC)
library(plotmo)
source("a_insights_shap_functions.r")
```

# Data Overview
```{r}
head(total_data, 5)
str(total_data)
dim(total_data) 
```

## Variable list

1. `Neo.Reference.ID` - the reference ID assigned to an asteroid
2. `Name` - the name given to an asteroid 
3. `Absolute.Magnitude` - the visual magnitude of an asteroid at zero phase angle and at unit heliocentric and geocentric distances
4. `Est.Dia.in.KM.min.` - the estimated minimum diameter of the asteroid in kilometers (KM) 
5. `Est.Dia.in.KM.max.` - the estimated maximum diameter of the asteroid in kilometers (KM)  
6. `Est.Dia.in.M.min.` - the estimated minimum diameter of the asteroid in meters (M)  
7. `Est.Dia.in.M.max.` - the estimated maximum diameter of the asteroid in meters (M) 
8. `Est.Dia.in.Miles.min.` - the estimated minimum diameter of the asteroid in miles    
9. `Est.Dia.in.Miles.max.` - the estimated maximum diameter of the asteroid in miles
10. `Est.Dia.in.Feet.min.` - the estimated minimum diameter of the asteroid in feet      
11. `Est.Dia.in.Feet.max.` - the estimated maximum diameter of the asteroid in miles          
12. `Close.Approach.Date` - the date of when the asteroid will be near Earth         
13. `Epoch.Date.Close.Approach` - the reference of time of when the asteroid will be near Earth
14. `Relative.Velocity.km.per.sec` - the relative velocity of the asteroid in kilometer per second
15. `Relative.Velocity.km.per.hr` - the relative velocity of the asteroid in kilometer per hour
16. `Miles.per.hour` - the velocity of the asteroid in kilometer per hour             
17. `Miss.Dist..Astronomical.` - the miss distance (measure proximity) of the asteroid in astronomical unit    
18. `Miss.Dist..lunar.` - the miss distance (measure proximity) of the asteroid in lunar distance      
19. `Miss.Dist..kilometers.` -  the miss distance (measure proximity) of the asteroid in kilometers     
20. `Miss.Dist..miles.` - the miss distance (measure proximity) of the asteroid in miles        
21. `Orbiting.Body` - the planet around which the asteroid is revolving                
22. `Orbit.ID` - the ID for the orbit                     
23. `Orbit.Determination.Date` - the date when the orbit is determined/found    
24. `Orbit.Uncertainity` - the uncertainty of the orbit          
25. `Minimum.Orbit.Intersection` - the potential close approaches and collision risks between astronomical objects   
26. `Jupiter.Tisserand.Invariant` - distinguish different kinds of orbits
27. `Epoch.Osculation` - the orbit that is calculated from the position and velocity vectors at a particular instant of time (i.e., epoch)            
28. `Eccentricity` - a measure of how far from circular each orbit is: the smaller the eccentricity number, the more circular the realm                
29. `Semi.Major.Axis` - the value of the longest diameter (i.e., the Semi Major Axis) of the asteroid’s orbit             
30. `Inclination` -  the angle between the orbit plane and the ecliptic plane                 
31. `Asc.Node.Longitude` - the angle in the ecliptic plane between the inertial-frame x-axis and the line through the ascending node.           
32. `Orbital.Period` - the time taken by the asteroid to make one full revolution around its orbiting body
33. `Perihelion.Distance` - the distance of the asteroid's closest point to the Sun           
34. `Perihelion.Arg` - the angle in the orbit plane between the ascending node and the perihelion point
35. `Aphelion.Dist` - the distance of the asteroid's farthest point to the Sun              
36. `Perihelion.Time` - the length of time to complete Perihelion (unit unknown)            
37. `Mean.Anomaly` - the mean of the angles in the orbit plane between the perihelion point and the position of the orbiting object                 
38. `Mean.Motion` - the angular speed required for a body to complete one orbit                 
39. `Equinox` - the moment in time when the Sun crosses the asteroid's equator                      
40. `Hazardous` - whether the asteroid is hazardous or not   

# Data Cleaning

Just based on the description/definition of each variable, some of them are identical and some captures the same value but in different units. For the sake of our analysis, we will drop those repeated columns:

```{r}
# Neo.Reference.ID and Name are identical
all(total_data$Neo.Reference.ID == total_data$Name)
```

`Neo.Reference.ID` and `Name` are identical columns, so `Name` will be kept. `Est.Dia.in.KM.min.`, `Est.Dia.in.M.min.`, `Est.Dia.in.Miles.min.`, `Est.Dia.in.Feet.min.` are the same value in different units. Same for the max. For the sake of our analysis, we will only use `Est.Dia.in.KM.min.` and `Est.Dia.in.KM.max.`. To maintain consistency in units, we will use `Relative.Velocity.km.per.hr`. For Miss.Dist., we will only use `Miss.Dist..Astronomical.`. Also, `Equinox` only has one value J2000 and `Orbiting.Body` only has Earth which probably are not good predictors. Since our data captures the status of 4687 asteroids at an Epoch (i.e., at a certain time), `Close.Approach.Date` and `Orbit.Determination.Date` may not be good reference for our current time frame and thus will be removed

```{r}
# Filter data
nasa_data <- total_data[, - c(1:2, 6:12, 14, 18:21, 23, 39)]

# Check Points
summary(nasa_data) # no NAs
str(nasa_data) # correct data format
nrow(unique(nasa_data)) == nrow(nasa_data) # no replicates
```

We will set `Hazardous` as our response variable and there are 755 cases of hazardous and 3932 cases of non-hazardous, which is imbalanced data

# Visualization

```{r, message=FALSE, warning=FALSE, out.width="200%"}
# Imbalanced Response Variable
ggplot(nasa_data, aes(x = Hazardous)) +
  geom_bar()+ 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        panel.background = element_blank(),
        text = element_text(size = 15)) 

# Boxplot all variables
m_dat <- melt(nasa_data, id.vars = "Hazardous") 

ggplot(m_dat, aes( y = value, x = factor(Hazardous), fill = factor(Hazardous))) + 
  geom_boxplot() + 
  facet_wrap(~ variable, scales = "free") + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        panel.background = element_blank(),
        text = element_text(size = 15)) + 
  labs(x = "Hazardous", title = "Hazardous or not", fill = "Hazardous") + 
  scale_fill_manual(values = c("True" = "red", "False" = "blue"), 
                    labels = c("True" = "Hazardous", "False" = "Not Hazardous"))

# Correlation matrix
nasa_corr <- cor(nasa_data[,-24])
ggcorrplot(nasa_corr, method = c('square'),hc.order = TRUE, type = "lower", outline.color = 'white') +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 90),
        text = element_text(size = 15)) + 
  labs(x = "", title = "Correlation Matrix", y = "") 
```

# Data Partition

```{r}
 set.seed(123456) 
 
# Perform stratified sampling
 split_data <- stratified(nasa_data, 
                         group = "Hazardous", 
                         size = 0.2,  # set for test 
                         bothSets = TRUE ) 
 # Extract train data
 train_data <- split_data[[2]]
 # Extract test data
 test_data <- split_data[[1]]

# Check size
nrow(train_data)
nrow(test_data)
```

# Trees 

## Random Forest 

```{r}
set.seed(123456)
rf1 <- randomForest(Hazardous ~., 
                data = train_data,
                ntree = 200, # Set number of trees
                nodesize = 200)  # Set node size

rf1_pred <- predict(rf1, test_data) # Create predictions 

t_rf <- table(rf1_pred,test_data$Hazardous) # Create table
rf_acc <- confusionMatrix(t_rf,  positive = "True") # Produce confusion matrix
rf_acc
```

Random Forest generates very high accuracy at 99.47%, sensitivity at 98.01% and specificity at 99.75%.

## XGBoost

### Convert Data 

```{r}
# Create training matrix
dtrain <- xgb.DMatrix(data = as.matrix(train_data[, 1:23]), label = as.numeric(train_data$Hazardous) -1)

# Create test matrix
dtest <- xgb.DMatrix(data = as.matrix(test_data[, 1:23]), label = as.numeric(test_data$Hazardous) - 1)
```

### Fit XGBoost Model

```{r}
set.seed(123456)
bst_1 <- xgboost(data = dtrain, # Set training data
               nrounds = 100, # Set number of rounds
               verbose = 1, # 1 - Prints out fit
               print_every_n = 20, # Prints out result every 20th iteration
               objective = "binary:logistic", # Set objective
               eval_metric = "auc",
               eval_metric = "error") # Set evaluation metric to use

# Prediction
boost_preds_1 <- predict(bst_1, dtest) # Create predictions for xgboost model
pred_dat <- cbind.data.frame(boost_preds_1 , test_data$Hazardous) # Issue
# Convert predictions to classes, using optimal cut-off
boost_pred_class <- rep("False", length(boost_preds_1))
boost_pred_class[boost_preds_1 >= 0.5] <- "True"

t_xgb <- table(boost_pred_class, test_data$Hazardous) # Create table
xgb_acc <- confusionMatrix(t_xgb, positive = "True") # Produce confusion matrix
xgb_acc
```

XGBoost generates accuracy at 99.57%, sensitivity at 98.68% and specificity at 99.75%.

### Imbalanced Data

```{r}
summary(train_data$Hazardous) # Imbalanced Data
zero_weight <- sum(train_data$Hazardous == 'False')/sum(train_data$Hazardous == 'True')

set.seed(123456)
bst_bal <- xgboost(data = dtrain, # Set training data
              eta = 0.05, # Set learning rate
              max.depth =  7, # Set max depth
              min_child_weight = 10, # Set minimum number of samples in node to split
              gamma = 0, # Set minimum loss reduction for split
              subsample =  0.9, # Set proportion of training data to use in tree
              colsample_bytree = 0.9, # Set number of variables to use in each tree
               
              nrounds = 200, # Set number of rounds
              early_stopping_rounds = 20, # Set number of rounds to stop at if there is no improvement
               
              verbose = 1, # 1 - Prints out fit
              nthread = 1, # Set number of parallel threads
              print_every_n = 20, # Prints out result every 20th iteration
              
              scale_pos_weight = zero_weight, # Set weights
              
              objective = "binary:logistic", # Set objective
              eval_metric = "auc",
              eval_metric = "error") # Set evaluation metric to use

# Prediction
boost_preds_bal <- predict(bst_bal, dtest) # Create predictions for xgboost model
pred_dat <- cbind.data.frame(boost_preds_bal , test_data$Hazardous) # Issue
# Convert predictions to classes, using optimal cut-off
boost_pred_class <- rep("False", length(boost_preds_bal))
boost_pred_class[boost_preds_bal >= 0.5] <- "True"

t_xgb_bal <- table(boost_pred_class, test_data$Hazardous) # Create table
xgb_ba_acc <- confusionMatrix(t_xgb_bal, positive = "True") # Produce confusion matrix
xgb_ba_acc
```

After adjusted the model for imbalanced data, the accuracy is at 99.15%, sensitivity at 99.34% and specificity at 99.11%., which doesn't differ too much.


### Variable Importance

```{r, message=FALSE}
x_vars <- model.matrix(Hazardous ~., data = train_data)[,-1]

shap_result <- shap.score.rank(xgb_model = bst_1, 
                X_train =x_vars,
                shap_approx = F)

shap_long = shap.prep(shap = shap_result,
                           X_train = x_vars, 
                           top_n = 10)

plot.shap.summary(data_long = shap_long)
```

Based on Shap calculation, `Absolute.Magnitude` and `Minimum.Orbit.Intersection` are quite important features. Let's try remove those two and run XGBoost again.

```{r}
# Set data again
dtrain1 <- xgb.DMatrix(data = as.matrix(train_data[, c(2:9, 11:23)]), 
                       label = as.numeric(train_data$Hazardous) -1)
dtest1 <- xgb.DMatrix(data = as.matrix(test_data[, c(2:9, 11:23)]), 
                     label = as.numeric(test_data$Hazardous) - 1)

# Run Model
set.seed(123456)
bst_2 <- xgboost(data = dtrain1, # Set training data
               nrounds = 100, # Set number of rounds
               verbose = 1, # 1 - Prints out fit
               print_every_n = 20, # Prints out result every 20th iteration
               objective = "binary:logistic", # Set objective
               eval_metric = "auc",
               eval_metric = "error") # Set evaluation metric to use

# Prediction
boost_preds_2 <- predict(bst_2, dtest1) # Create predictions for xgboost model
pred_dat <- cbind.data.frame(boost_preds_2 , test_data$Hazardous) # Issue
# Convert predictions to classes, using optimal cut-off
boost_pred_class <- rep("False", length(boost_preds_2))
boost_pred_class[boost_preds_2 >= 0.5] <- "True"

t_xgb_2 <- table(boost_pred_class, test_data$Hazardous) # Create table
confusionMatrix(t_xgb_2, positive = "True") # Produce confusion matrix
```

Now, `bst_2` gives us the accuracy at 94.24%, sensitivity at 72.85% and specificity at 98.35%.

```{r}
shap_result <- shap.score.rank(xgb_model = bst_2, 
                X_train =as.matrix(train_data[, c(2:9, 11:23)]),
                shap_approx = F)

shap_long = shap.prep(shap = shap_result,
                           X_train = as.matrix(train_data[, c(2:9, 11:23)]), 
                           top_n = 10)

plot.shap.summary(data_long = shap_long)
```

# Logistic Regression

## Fitting Logistic Regression

```{r, warning=FALSE, message=FALSE}
glm_full <- glm(Hazardous ~ ., data = train_data, family = 'binomial')

# Variable Selection Using BIC Standard
glm_bwd <- step(glm_full, direction = 'backward', k = log(nrow(nasa_data)))
summary(glm_bwd)
```

## Apply LASSO

```{r}
x_vars <- as.data.frame(scale(nasa_data[, -24]))

# Create sequence of lambda values
lambda_seq <- seq(from = 0.1, to = 10, by = 0.1)
# Run lasso cross validation
glm_lasso_cv <- cv.glmnet(x = as.matrix(x_vars), # Set explanatory variables
                   y = as.numeric(nasa_data$Hazardous), # Set response variable
                   alpha = 1, # Set alpha as 1 for lasso
                   lambda = lambda_seq, # Set lambda as sequence of lambda values
                   nfolds = 10) # Set number of folds as 10

best_lam <- glm_lasso_cv$lambda.1se # Extract best lambda
best_lam # View best lambda

glm_lasso <- glmnet(x = x_vars,
                    y = nasa_data$Hazardous,
                    alpha = 1, 
                    family = "binomial", 
                    lambda = best_lam)

coef(glm_lasso)
```

## Model Evaluation

```{r, warning=FALSE}
glm_full_pred <- predict(glm_full, newdata = test_data, type = 'response')
glm_bwd_pred <- predict(glm_bwd, newdata = test_data, type = 'response')

glm_full_acc <- confusionMatrix(factor(ifelse(glm_full_pred > 0.5, 'True', 'False')), test_data$Hazardous, positive = 'True')
print(glm_full_acc)

glm_bwd_acc <- confusionMatrix(factor(ifelse(glm_bwd_pred > 0.5, 'True', 'False')), test_data$Hazardous, positive = 'True')
print(glm_bwd_acc)
```

# Neural Network Model

## Standardization
```{r}
# Scale train data
x_train_nn <- model.matrix(~ Absolute.Magnitude + Est.Dia.in.KM.min. + Est.Dia.in.KM.max. +
                             Epoch.Date.Close.Approach + Relative.Velocity.km.per.hr +
                             Miles.per.hour + Miss.Dist..Astronomical. + Orbit.ID +
                             Orbit.Uncertainity + Minimum.Orbit.Intersection +
                             Jupiter.Tisserand.Invariant + Epoch.Osculation + Eccentricity +
                             Semi.Major.Axis + Inclination + Asc.Node.Longitude +
                             Orbital.Period + Perihelion.Distance + Perihelion.Arg +
                             Aphelion.Dist + Perihelion.Time + Mean.Anomaly + Mean.Motion,
                           data = train_data)[,-1]
  
# standardization
x_mean <- apply(x_train_nn, 2, mean)
x_sd <- apply(x_train_nn, 2, sd)
x_train_nn <- scale(x_train_nn, center = x_mean, scale = x_sd)

# combine with dependent variable Outcome
x_train_nn <- cbind.data.frame(train_data$Hazardous, x_train_nn)
colnames(x_train_nn)[1] <- 'Hazardous'


# Scale test data
x_test_nn <- model.matrix(~ Absolute.Magnitude + Est.Dia.in.KM.min. + Est.Dia.in.KM.max. +
                             Epoch.Date.Close.Approach + Relative.Velocity.km.per.hr +
                             Miles.per.hour + Miss.Dist..Astronomical. + Orbit.ID +
                             Orbit.Uncertainity + Minimum.Orbit.Intersection +
                             Jupiter.Tisserand.Invariant + Epoch.Osculation + Eccentricity +
                             Semi.Major.Axis + Inclination + Asc.Node.Longitude +
                             Orbital.Period + Perihelion.Distance + Perihelion.Arg +
                             Aphelion.Dist + Perihelion.Time + Mean.Anomaly + Mean.Motion,
                           data = test_data)[,-1]

x_test_nn <- scale(x_test_nn, center = x_mean, scale = x_sd)

x_test_nn <- cbind.data.frame(test_data$Hazardous, x_test_nn)
colnames(x_test_nn)[1] <- 'Hazardous'
```

## Neural Network Model Creation

```{r, eval=FALSE}
# Decide the number of hidden layer and neurons
seq_1 <- seq(2, 10, 1)
params <- expand.grid(seq_1, seq_1)
accuracy_res <- rep(NA, nrow(params))
for (i in 56:nrow(params)){
  set.seed(123456)
  model <- neuralnet(Hazardous == 'True' ~ ., data = x_train_nn,
                     hidden = c(params[i,1],params[i,2]), linear.output = F)
  
  accuracy_res[i] <- confusionMatrix(factor(ifelse(predict(model, newdata = x_test_nn)[, 1] > 0.5, 'True', 'False')), test_data$Hazardous, positive = 'True')[3][[1]][['Accuracy']]
  
}

accuracy_table <- cbind(params, accuracy_res)

accuracy_table[which.max(accuracy_table$accuracy_res), ] # two layers with 5 neurons on 1st and 4 neurons on 2nd
```

```{r}
set.seed(123456)
nn1 <- neuralnet(Hazardous == 'True' ~ ., data = x_train_nn, hidden = c(5, 4), linear.output = F)
nn1_pred <- predict(nn1, newdata = x_test_nn)[, 1]

nn1_acc <- confusionMatrix(factor(ifelse(nn1_pred > 0.5, 'True', 'False')),
                           test_data$Hazardous, positive = 'True')
nn1_acc
```

# Model Comparison

```{r, message=FALSE}
# Compare Confusion Matrix

# Check accuracy
c(paste('XGBoost', round(xgb_acc[3]$overall[1],4), sep = ": "),
  paste('glm_bwd', round(glm_bwd_acc[3]$overall[1],4), sep = ": "),
  paste('nn', round(nn1_acc[3]$overall[1],4), sep = ": "))

# Check Sensitivity
c(paste('XGBoost', round(xgb_acc[4]$byClass[1],4), sep = ": "),
  paste('glm_bwd', round(glm_bwd_acc[4]$byClass[1],4), sep = ": "),
  paste('nn', round(nn1_acc[4]$byClass[1],4), sep = ": "))

# Check Specificity
c(paste('XGBoost', round(xgb_acc[4]$byClass[2],4), sep = ": "),
  paste('glm_bwd', round(glm_bwd_acc[4]$byClass[2],4), sep = ": "),
  paste('nn', round(nn1_acc[4]$byClass[2],4), sep = ": "))


# Plot ROC

# Calculate XGBoost ROC
roc1 = roc(test_data$Hazardous, boost_preds_1)
# Calculate Logistic Regression ROC
roc2 = roc(test_data$Hazardous, glm_full_pred)
# Calculate Neural Network ROC
roc3 = roc(test_data$Hazardous, nn1_pred)

plot.roc(roc1, print.auc = TRUE, print.auc.x = 0, print.auc.y = 0.4, col = "navy", print.auc.col = "navy")

plot.roc(roc2, print.auc = TRUE, print.auc.x = 0, print.auc.y = 0.6, col = "green", print.auc.col = "green", add = TRUE)

plot.roc(roc3, print.auc = TRUE, print.auc.x = 0, print.auc.y = 0.8, col = "orange", print.auc.col = "orange", add = TRUE)
```

